Verification of a TCP/IP stack

### 2021-05-18 SecDev21/paper.tex
minor fixes

\def\todo#1{\textcolor{red}{[#1]}}

@@

In this paper, we take a practical approach, where the security of an existing TCP/IP stack implementation written in ANSI C programming language is incrementally enhanced by replacing parts of its code with formally verified code in SPARK \cite{mccormick_chapin_2015}. SPARK 2014 is a programming language designed as a subset of the Ada language to produce highly reliable software through formal verification. For this work, the CycloneTCP, a professional-grade embedded TCP/IP library developed by the Oryx Embedded company \cite{CycloneTCP}. \Cref{Fig:TcpStack} shows the protocols supported by the library. This work focuses on the TCP and socket components of the stack. The library's TCP implementation is meant to conform with the RFC 793 protocol specifications \cite{rfc793}. The quality of the CycloneTCP is acknowledged by the AMNESIA:33 report \cite{AMNESIA33}, which classifies it as one of the most resilient TCP/IP stack. By using as a starting point such a well-tested and widely used TCP/IP implementation, the IoT industry can have an immediate benefit by adopting this work's security-enhanced version of the CycloneTCP library.

We present the TCP protocol in section~\ref{sec:TCP} and then briefly the
chosen stack, the Oryx Embedded one, in section~\ref{sec:stack}
before presenting the specification techniques used in section~\ref{sec:spec}
and then in sections~\ref{sec:verif} and~\ref{sec:API} we explain how we
formally implemented in SPARK the TCP user functions and how we hardened the
user's API. Finally we end in section~\ref{sec:results} by showing the results
of our work.

@@

The CycloneTCP implementation adopts the three-task-based model and the socket data structure. Thus, one task is dedicated to the user's code, one to the timers, and one to the processing of the incoming segments. The user can manipulate a socket through the TCP's library Application Programming Interface (API) to perform various TCP operations, such as data transmission.

@@

SPARK~\cite{mccormick2015building} is the name of a platform that provides
formal verification for Ada. It uses the user-provided contracts and attempts
to prove that the runtime checks cannot fail and that postconditions are
established by the corresponding subprograms.  As formal verification for the
whole Ada language would be intractable, SPARK is also the name of the subset
of the Ada language that is supported by the SPARK tool called
GNATprove\footnote{\url{http://docs.adacore.com/spark2014-docs/html/ug/}}.
This subset contains almost all features of Ada, though sometimes in a
restricted form.  In particular, expressions should be free from side effects,
and aliasing is forbidden (no two variables should share the same memory
location or overlap in memory), including when using pointers thanks to the use
of an ownership policy~\cite{dross2020recursive}.  This restriction greatly
simplifies the memory model used in the SPARK tool: any program variables can
be reasoned about independently from other variables.

>>>>

\usepackage{todonotes}
%\def\todo#1{\textcolor{red}{[#1]}}

@@

In this paper, we take a practical approach, where the security of an existing TCP/IP stack implementation written in ANSI C programming language is incrementally enhanced by replacing parts of its code with formally verified code in SPARK \cite{mccormick_chapin_2015}. SPARK 2014 is a programming language designed as a subset of the Ada language to produce highly reliable software through formal verification. For this work, the CycloneTCP, a professional-grade embedded TCP/IP library developed by the Oryx Embedded company \cite{CycloneTCP}. \Cref{Fig:TcpStack} shows the protocols supported by the library. This work focuses on the TCP and socket components of the stack. The library's TCP implementation is meant to conform with the RFC 793 protocol specifications \cite{rfc793}. The quality of the CycloneTCP is acknowledged by the AMNESIA:33 report \cite{AMNESIA33}, which classifies it as one of the most resilient TCP/IP stack. By using as a starting point such a well-tested and widely used TCP/IP implementation, the IoT industry can have an immediate benefit by adopting this work's security-enhanced version of the CycloneTCP library. 
\todo[inline]{pending contributions list and fixing the next paragraph.}
We present the TCP protocol in section~\ref{sec:TCP} and then briefly the
chosen stack, the Oryx Embedded one, in section~\ref{sec:stack}
before presenting the specification techniques used in section~\ref{sec:spec}
and then in sections~\ref{sec:verif} and~\ref{sec:API} we explain how we
formally implemented in SPARK the TCP user functions and how we hardened the
user's API. Finally we end in section~\ref{sec:results} by showing the results
of our work.

@@

The CycloneTCP implementation adopts the three-task-based model and the socket data structure defined by the \cite{rfc793}. Thus, it has tasks dedicated to the user's code, to the timers, and to the processing of the incoming segments. Moreover, the user can manipulate a socket through the TCP's library Application Programming Interface (API) to perform various TCP operations, such as data transmission.

@@

SPARK~\cite{mccormick_chapin_2015} is the name of a platform that provides
formal verification for Ada. It uses the user-provided contracts and attempts
to prove that the runtime checks cannot fail and that postconditions are
established by the corresponding subprograms.  As formal verification for the
whole Ada language would be intractable, SPARK is also the name of the subset
of the Ada language that is supported by the SPARK tool called
GNATprove\footnote{\url{http://docs.adacore.com/spark2014-docs/html/ug/}}.
This subset contains almost all features of Ada, though sometimes in a
restricted form.  In particular, expressions should be free from side effects,
and aliasing is forbidden (no two variables should share the same memory
location or overlap in memory), including when using pointers thanks to the use
of an ownership policy~\cite{dross2020recursive}.  This restriction greatly
simplifies the memory model used in the SPARK tool: any program variables can
be reasoned about independently from other variables.
